---
title: "Subclustering"
output:
  html_document:
    code_download: yes
    code_folding: hide
    css: ../"style.css"
    dev: "png"
    df_print: kable
    fig_caption: yes
    keep_md: no
    self_contained: true
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
        collapsed: false
        smooth_scroll: false
editor_options:
  chunk_output_type: console
---

------------------------------------------------------------------------

# Overview and setup

```{r setup-options, include = FALSE}
knitr::opts_chunk$set(comment = NA, dpi = 200, fig.align = "center",
                      out.width = "100%", warning = FALSE, error = TRUE,
                      message = FALSE,
                      fig.path = here::here("primary_analysis", "figures",
                                            "analysis", "subclustering",
                                            "subclustering-"))

# will mess with chunk labels
options(knitr.duplicate.label = "allow")
```

**Author(s):** Edel Aron<br>
**Date Created:** 2023-02-22<br>
**Last Updated:** `r Sys.Date()`

**R version:** `r R.Version()$version.string`<br>
**Platform:** `r R.Version()$platform`<br>
**Running under:** `r sessionInfo()$running`

```{r, include = FALSE}
# load packages, colors and themes and functions (R code only)
code_setup <-
  knitr::purl(here::here("single_cell", "primary_analysis", "notebooks",
                         "01-setup.Rmd"),
              output = tempfile(), documentation = 0)
code_functions <-
  knitr::purl(here::here("single_cell", "primary_analysis", "notebooks",
                         "02-functions.Rmd"),
              output = tempfile(), documentation = 0)
code_meta <- 
  knitr::purl(here::here("single_cell", "primary_analysis", "notebooks",
                         "03-meta.Rmd"),
              output = tempfile(), documentation = 0)

# read in the code, then remove the temp files
source(code_setup)
source(code_functions)
source(code_meta)
unlink(c(code_setup, code_functions, code_meta))
```

```{r}
include_graphics(path = here::here("single_cell", "primary_analysis", "images",
                                   "single_cell_subclustering_pipeline.png"))
```

------------------------------------------------------------------------

# Read in the data

## Markers

```{r setup-markers}
markers_all <- read_csv(file.path(path_resources, "marker_genes_all.csv"),
                        show_col_types = FALSE)
```

## Objects

### GEX

```{r setup-gex}
# objects
combined_blood <-
  readRDS(file.path(path_objs, "airr", "combined_blood_airr.rds"))
combined_skin <-
  readRDS(file.path(path_objs, "airr", "combined_skin_airr.rds"))

# colors
colors_annotated_blood <- combined_blood@misc$colors_annotated
colors_annotated_skin <- combined_skin@misc$colors_annotated
```

Just like in the GEX analysis section, you can load in the processed objects instead of running the pipeline if it was previously run with the same settings. It is recommended that you save the objects after an initial (or new) analysis since rerunning the previous steps is time-intensive (and liable to sometimes crash the session depending on the configuration).

```{r setup-gex-subclustered}
# objects
combined_blood_t <- readRDS(file.path(path_objs, "combined_blood_t.rds"))
combined_skin_t <- readRDS(file.path(path_objs, "combined_skin_t.rds"))

# colors
colors_annotated_blood_t <- combined_blood_t@misc$colors_annotated
colors_annotated_skin_t <- combined_skin_t@misc$colors_annotated
```

```{r setup-gex-subclustered-overview}
subcluster_overview(seurat_obj = combined_blood,
                    seurat_obj_sub = combined_blood_t,
                    tissue_type = "Blood", airr_type = "TCR",
                    cell_types = t_cell_types)
subcluster_overview(seurat_obj = combined_skin,
                    seurat_obj_sub = combined_skin_t,
                    tissue_type = "Skin", airr_type = "TCR",
                    cell_types = t_cell_types)
```

Reset to the unannotated clusters:

```{r setup-gex-subclustered-labels}
# since later I save the objects with their annotations
Idents(combined_blood_t) <- "seurat_subclusters"
Idents(combined_skin_t) <- "seurat_subclusters"
```

### AIRR

```{r setup-airr}
combined_tcr <- readRDS(file.path(path_airrflow_objs,
                                  "combined_airrflow_tcr.rds"))
```

------------------------------------------------------------------------

# Blood

## Object setup

### Select the clusters

```{r blood-t-setup-select}
#| eval = FALSE

combined_blood_t <- subset(combined_blood,
                             Has_TCR == TRUE | # AIRR info
                             annotated_clusters %in% t_cell_types) # GEX info
```

```{r blood-t-setup-overview}
subcluster_overview(seurat_obj = combined_blood,
                    seurat_obj_sub = combined_blood_t,
                    tissue_type = "Blood", airr_type = "TCR",
                    cell_types = t_cell_types)
```

### Rerun the standard pipeline

Choose a resolution:

```{r blood-t-seurat-pipeline-res}
blood_t_res <- 0.6
```

With the chosen resolution (with help from `clustree`) of `r blood_t_res`:

```{r blood-t-seurat-pipeline-save}
#| eval = FALSE

# resolution picked from the clustering tree, plus an extra one
combined_blood_t <-
  seurat_pipeline(seurat_obj = combined_blood_t, subcluster = TRUE,
                  num_features = 2000, num_pcs = 30, num_dims = 20, 
                  cluster_res = c(blood_t_res, 1.1))

# choose a resolution
Idents(combined_blood_t) <- paste0("RNA_snn_sub_res.", blood_t_res)
combined_blood_t$seurat_subclusters <-
  combined_blood_t[[paste0("RNA_snn_sub_res.", blood_t_res)]]

# save the object
saveRDS(combined_blood_t, file.path(path_objs, "combined_blood_t.rds"))
```

### Examine results

```{r blood-t-seurat-pipeline-results-elbow-bar}
#| fig.dim = c(10, 12)

# plot variable features with labels for the top 10
p1 <- LabelPoints(VariableFeaturePlot(combined_blood_t),
                  points = head(VariableFeatures(combined_blood_t), 10),
                  xnudge = 0, ynudge = 0) +
        labels_standard + legend_bottom

p2 <- ElbowPlot(combined_blood_t, ndims = 25) + labels_standard

p3 <- plot_counts_cluster(clusters = combined_blood_t$seurat_subclusters,
                          tissue_type = "Blood",
                          details = "Subclustered T Cells",
                          x_axis = "Cluster")

((p1 | p2) / p3) + plot_layout(heights = c(2, 3))
```

## Dimensionality reduction {.tabset}

### Clusters by Samples

```{r blood-t-clusters-samples-bar}
#| fig.dim = c(16, 15)

# percentage of clusters per sample
p1 <- plot_pcts(pcts = calc_pcts(data = combined_blood_t[[]],
                                 meta_group_by = c("SampleName", "Dataset"),
                                 focus_group = "seurat_subclusters"),
                tissue_type = "Blood", plot_value = "Clusters",
                x_axis = "SampleName", x_axis_label = "Sample",
                fill_type = "seurat_subclusters", fill_label = "Cluster",
                plot_type = "All",
                details = t_cell_types_str) %>%
        add_info_bar(method = "contains", info_type = "Dataset")

# see which samples are going into each cluster
p2 <- plot_tile(seurat_obj = combined_blood_t, tissue_type = "Blood",
                clrs_specific = colors_sample_blood,
                x_axis = "seurat_subclusters", x_axis_label = "Seurat Cluster",
                details = t_cell_types_str) %>%
        add_info_bar(info_type = "Dataset", top_side = FALSE)

# percentage of samples per cluster
p3 <- plot_pcts(pcts = calc_pcts(data = combined_blood_t[[]],
                                 meta_group_by = "seurat_subclusters",
                                 focus_group = "SampleName"),
                tissue_type = "Blood", plot_value = "Samples",
                x_axis = "seurat_subclusters", x_axis_label = "Cluster",
                fill_type = "SampleName", fill_label = "Sample",
                clrs_specific = colors_sample_blood,
                plot_type = "All", details = t_cell_types_str)

# percentage of clusters per dataset
p4 <- plot_pcts(pcts = calc_pcts(data = combined_blood_t[[]],
                                 meta_group_by = "Dataset",
                                 focus_group = "seurat_subclusters"),
                tissue_type = "Blood", plot_value = "Clusters",
                x_axis = "Dataset", x_axis_label = "Dataset",
                fill_type = "seurat_subclusters", fill_label = "Cluster",
                details = t_cell_types_str) %>%
        add_info_bar(method = "contains", info_type = "Dataset")

# percentage of datasets per cluster
p5 <- plot_pcts(pcts = calc_pcts(data = combined_blood_t[[]],
                                 meta_group_by = "seurat_subclusters",
                                 focus_group = "Dataset"),
                tissue_type = "Blood", plot_value = "Datasets",
                x_axis = "seurat_subclusters", x_axis_label = "Cluster",
                fill_type = "Dataset", clrs_specific = colors_dataset,
                details = t_cell_types_str)

# combine all plots
((p1 / p4 + plot_layout(guides = "collect") |
((p2 / p3 / p5) + plot_layout(guides = "collect"))) &
    guides(fill = guide_legend(ncol = 1)))
```

### Cells by Samples

```{r blood-t-cells-samples-bar}
#| fig.dim = c(12, 8)

plot_counts_cluster(clusters = combined_blood_t$SampleName,
                    tissue_type = "Skin",
                    details = "Subclustered T Cells",
                    x_axis = "Cell Type",
                    clrs_specific = colors_samples) +
  labels_rotate_x
```

### Cell Types by Samples

```{r blood-t-cell-types-samples-bar}
#| fig.dim = c(12, 8)

plot_counts_cluster(clusters = combined_blood_t$annotated_clusters,
                    tissue_type = "Skin",
                    details = "Subclustered T Cells",
                    x_axis = "Cell Type",
                    clrs_specific = colors_annotated_blood) +
  labels_rotate_x
```

### Samples

```{r blood-t-samples-bar}
#| fig.dim = c(12, 7)

plot_pcts(calc_pcts(data = combined_blood_t[[]] %>%
                      mutate(Type = "Sample"),
                    meta_group_by = "Type",
                    focus_group = "SampleName"),
          tissue_type = "Subclustered Blood T Cells",
          clrs_specific = colors_sample_blood,
          plot_value = "Sample", x_axis = "Type", x_axis_label = "Total",
          fill_type = "SampleName", fill_label = "Sample",
          perc_min = 2, label_size = 3, label_fill = TRUE,
          include_counts = FALSE) +
  coord_flip() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(ncol = 1))
```

### UMAPs

All:

```{r blood-t-umaps}
#| fig.dim = c(24, 8)

p1 <- plot_umap(seurat_obj = combined_blood_t,
                tissue_type = "Subclustered Blood T Cells",
                use_hues = TRUE, plot_by = "cluster_all",
                clusters_col = "seurat_subclusters")
p2 <- plot_umap(seurat_obj = combined_blood_t,
                tissue_type = "Subclustered Blood T Cells",
                clrs_specific = colors_sample_blood,
                plot_by = "all", plot_label = FALSE,
                clusters_col = "seurat_subclusters")
p3 <- plot_umap(seurat_obj = combined_blood_t,
                tissue_type = "Subclustered Blood T Cells",
                clrs_specific = colors_dataset,
                plot_by = "dataset", plot_label = FALSE,
                clusters_col = "seurat_subclusters")

p1 | p2 | p3
```

By Sample:

```{r blood-t-umap-cluster-sample}
#| fig.dim = c(16, 8)

plot_umap(seurat_obj = combined_blood_t,
          tissue_type = "Subclustered Blood T Cells",
          use_hues = TRUE, plot_by = "cluster_sample",
          clusters_col = "seurat_subclusters")
```

## Marker overlays

### Setup

```{r blood-t-marker-overlays-setup}
#| class.source = "fold-show"

# all T cell markers
blood_t_marker_types <- c("NK", "T_CD4", "T_CD4RM", "T_CD8", "T_CD8RM",
                          "T_Dividing", "T_FH", "T_Reg", "TCM",
                          "TCR", "TEM", "TRM")

get_features_from_all(markers_df = markers_all, cell_types = blood_t_marker_types)
```

### Dot plot

```{r blood-t-marker-overlays-dot}
#| fig.dim = c(18, 6)

DotPlot(combined_blood_t,
        features = get_features_from_all(markers_df = markers_all,
                                         cell_types = blood_t_marker_types,
                                         alphabetize_all = FALSE),
        cols = colors_dot, dot.scale = 3, cluster.idents = TRUE) %>%
  add_info_bar(method = "join", info_type = "Cell_Type_Full",
               info = get_cell_types(markers_all %>%
                                       filter(Cell_Type %in% blood_t_marker_types)),
               text_size = 6, label = FALSE, angle = 90) %>%
  plot_dot_side(seurat_obj = combined_blood_t,
                row_identity = "seurat_subclusters",
                facet_col = "Cell_Type_Full",
                info_to_add = c("cluster_size", "percent_TCR")) +
  labs(title = "Subclustered Blood T Cells") +
  RotatedAxis() + labels_standard
```

### Feature plots {.tabset}

#### CD4 T Cells

```{r blood-t-marker-overlays-feature-cd4}
#| fig.dim = c(16, 16)

FeaturePlot(combined_blood_t,
            features = get_features_from_all(markers_df = markers_all,
                                             cell_types = c("T_CD4", "T_CD4RM")),
            min.cutoff = 0,
            label = TRUE, label.size = 2, ncol = 4, raster = FALSE) +
  plot_annotation(title = "Subclustered Blood - CD4 T Cells") &
  labels_standard & clean_umap
```

#### CD8 T Cells

```{r blood-t-marker-overlays-feature-cd8}
#| fig.dim = c(16, 16)

FeaturePlot(combined_blood_t,
            features = get_features_from_all(markers_df = markers_all,
                                             cell_types = c("T_CD8", "T_CD8RM")),
            min.cutoff = 0,
            label = TRUE, label.size = 2, ncol = 4, raster = FALSE) +
  plot_annotation(title = "Subclustered Blood - CD8 T Cells") &
  labels_standard & clean_umap
```

#### Other T Cells

```{r blood-t-marker-overlays-feature-other}
#| fig.dim = c(16, 32)

FeaturePlot(combined_blood_t,
            features = get_features_from_all(markers_df = markers_all,
                                             cell_types = c("T_Dividing", "T_FH",
                                                            "T_Reg", "TCM",
                                                            "TEM", "TRM")),
            min.cutoff = 0,
            label = TRUE, label.size = 2, ncol = 4, raster = FALSE) +
  plot_annotation(title = "Subclustered Blood - Other T Cells") &
  labels_standard & clean_umap
```

#### TCRs

```{r blood-t-marker-overlays-feature-tcr}
#| fig.dim = c(16, 8)

FeaturePlot(combined_blood_t,
            features = get_features_from_all(markers_df = markers_all,
                                             cell_types = "TCR"),
            min.cutoff = 0,
            label = TRUE, label.size = 2, ncol = 4, raster = FALSE) +
  plot_annotation(title = "Subclustered Blood - TCRs") &
  labels_standard & clean_umap
```

## Annotate cell clusters

### Manually

```{r blood-t-annotations-manual}
#| eval = FALSE

combined_blood_t_annotations <- rbind(c("0", "TBD T Cells"),
                                      c("1", "TBD T Cells"),
                                      c("2", "TBD T Cells"),
                                      c("3", "TBD T Cells"),
                                      c("4", "TBD T Cells"), # NK T cells?
                                      c("5", "TBD T Cells"),
                                      c("6", "CD8+ T Cells"),
                                      c("7", "CD4+ T Cells"), # CD4
                                      c("8", "CD4+ T Cells"), # CD4, regulatory?
                                      c("9", "TBD T Cells"),
                                      c("10", "CD8+ T Cells"), # CD8, NK T cells?
                                      c("11", "TBD T Cells"),
                                      c("12", "CD8+ T Cells"), # CD8
                                      c("13", "TBD T Cells"),
                                      c("14", "CD8+ T Cells"), # CD8
                                      c("15", "CD8+ T Cells"), # CD8
                                      c("16", "TBD T Cells"),
                                      c("17", "Cycling T Cells"), # MKI67, TOP2A
                                      c("18", "Non T Cells")) # monocytes
colnames(combined_blood_t_annotations) <- c("Cluster", "CellType")
combined_blood_t_annotations <- data.frame(combined_blood_t_annotations)

# save the annotations as a csv
write_csv(combined_blood_t_annotations,
          file.path(path_tables, "annotations_blood_t.csv"))
```

Add the annotations to the Seurat object and examine them:

```{r blood-t-annotations-manual-add}
# read in and show the annotations
annotations_blood_t <- read_csv(file.path(path_tables,
                                          "annotations_blood_t.csv"),
                                col_types = "c")

# add the annotations to the object
combined_blood_t <- add_annotations(seurat_obj = combined_blood_t,
                                    annotations_df = annotations_blood_t,
                                    clusters_col = "seurat_subclusters",
                                    annotations_col = "annotated_subclusters")

# clusters by annotation
print_kable(annotations_blood_t %>%
              group_by(CellType) %>%
              transmute(Clusters = paste0(Cluster, collapse = ", ")) %>%
              distinct() %>%
              arrange(CellType),
            kable_height = NULL)
```

Choose the color scheme:

```{r blood-t-annotations-color-schemes}
# only want the non T cells to be black
blood_t_cell_types_num <- n_distinct(combined_blood_t$annotated_subclusters)
blood_t_cell_types <- sort(unique(combined_blood_t$annotated_subclusters))

colors_annotated_blood_t <-
    setNames(rep("#000000", blood_t_cell_types_num),
             nm = blood_t_cell_types)
colors_annotated_blood_t[str_subset(blood_t_cell_types, "Non", negate = TRUE)] <-
  iwanthue(n = blood_t_cell_types_num - 1, hmin = 250, hmax = 360)

# save the colors to the miscellaneous slot for easy access
Misc(combined_blood_t, slot = "colors_annotated") <- colors_annotated_blood_t
```

Save the annotated object:

```{r blood-t-annotations-manual-save}
#| eval = FALSE

# remove unneeded columns from the metadata
for (col in c("sequence_id_unique", "v_call_family_TRA", "j_call_family_TRA")) {
  combined_blood_t[[col]] <- c()
}

# save the object with annotations (watch out for reproducibility)
saveRDS(combined_blood_t, file.path(path_objs, "combined_blood_t.rds"))
```

### FindMarkers

All clusters:

```{r blood-t-annotations-findallmarkers-save}
#| eval = FALSE

Idents(combined_blood_t) <- "seurat_subclusters"

# find markers for every cluster compared to all remaining cells
# report only the positive ones
combined_blood_t_findallmarkers <- FindAllMarkers(combined_blood_t,
                                                  only.pos = TRUE, min.pct = 0.25,
                                                  logfc.threshold = 0.25)

Idents(combined_blood_t) <- "annotated_subclusters"

# save the markers as a csv
write_csv(combined_blood_t_findallmarkers,
          file.path(path_tables, "findmarkers",
                    "annotations_blood_t_findallmarkers.csv"))
```

```{r blood-t-annotations-findallmarkers-load}
# load markers
combined_blood_t_findallmarkers <-
  read_csv(file.path(path_tables, "findmarkers",
                     "annotations_blood_t_findallmarkers.csv"),
           show_col_types = FALSE)

print_kable(combined_blood_t_findallmarkers %>%
              group_by(cluster) %>%
              slice_max(n = 5, order_by = desc(p_val_adj)) %>%
              relocate(c("cluster", "gene", "p_val_adj"), .before = c("p_val")))
```

### Final annotations

```{r blood-t-annotations-manual-cluster-all-umap}
#| fig.dim = c(24, 8)

p1 <- plot_umap(seurat_obj = combined_blood_t,
                tissue_type = "Subclustered Blood T Cells",
                plot_by = "cluster_all",
                clusters_col = "seurat_subclusters") # for comparison
p2 <- plot_umap(seurat_obj = combined_blood_t,
                tissue_type = "Subclustered Blood T Cells",
                clrs_specific = colors_annotated_blood,
                plot_by = "cluster_all",
                annotated = TRUE, annotations_col = "annotated_clusters",
                details = "Blood Cell Type")
p3 <- plot_umap(seurat_obj = combined_blood_t,
                tissue_type = "Subclustered Blood T Cells",
                clrs_specific = colors_annotated_blood_t,
                plot_by = "cluster_all",
                annotated = TRUE, annotations_col = "annotated_subclusters")

p1 | p2 | p3
```

```{r blood-t-annotations-manual-cluster-sample-umap}
#| fig.dim = c(16, 8)

plot_umap(seurat_obj = combined_blood_t,
          tissue_type = "Subclustered Blood T Cells",
          clrs_specific = colors_annotated_blood_t,
          plot_by = "cluster_sample",
          annotated = TRUE, annotations_col = "annotated_subclusters")
```

## Check the annotations

### Cell counts per cluster

```{r blood-t-check-annotations-cell-type-counts-bar}
#| fig.dim = c(12, 8)

plot_counts_cluster(clusters = combined_blood_t$annotated_subclusters,
                    tissue_type = "Blood", details = "Subclustered T Cells",
                    x_axis = "Cell Type",
                    clrs_specific = colors_annotated_blood_t)
```

```{r blood-t-check-annotations-cell-type-counts-stacked-bar}
#| fig.dim = c(15, 5)

plot_pcts(calc_pcts(data = combined_blood_t[[]] %>% mutate(Type = "Sample"),
                    meta_group_by = "Type",
                    focus_group = "annotated_subclusters"),
          tissue_type = "Blood T Cells",
          clrs_specific = colors_annotated_blood_t,
          plot_value = "Cell Types", x_axis = "Type", x_axis_label = "Total",
          fill_type = "annotated_subclusters", fill_label = "Cell Type",
          perc_min = 1.5, label_size = 3, label_fill = TRUE,
          include_counts = FALSE, reverse_order = TRUE) +
  coord_flip() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(ncol = 1))
```

### Cell Types by Samples

```{r blood-t-check-annotations-cell-types-samples-bar}
#| fig.dim = c(16, 15)

# percentage of cell types per sample
p1 <- plot_pcts(pcts = calc_pcts(data = combined_blood_t[[]],
                                 meta_group_by = c("SampleName", "Dataset"),
                                 focus_group = "annotated_subclusters"),
                tissue_type = "Blood", plot_value = "Cell Types",
                x_axis = "SampleName", x_axis_label = "Sample",
                fill_type = "annotated_subclusters", fill_label = "Cell Type",
                clrs_specific = colors_annotated_blood_t,
                plot_type = "All", details = t_cell_types_str) %>%
        add_info_bar(method = "contains", info_type = "Dataset")

# see which samples are going into each cell type
p2 <- plot_tile(seurat_obj = combined_blood_t, tissue_type = "Blood",
                clrs_specific = colors_sample_blood,
                x_axis = "annotated_subclusters", x_axis_label = "Cell Type",
                details = t_cell_types_str) %>%
        add_info_bar(info_type = "Dataset", top_side = FALSE)

# percentage of samples per cell type
p3 <- plot_pcts(pcts = calc_pcts(data = combined_blood_t[[]],
                                 meta_group_by = "annotated_subclusters",
                                 focus_group = "SampleName"),
                tissue_type = "Blood", plot_value = "Samples",
                x_axis = "annotated_subclusters", x_axis_label = "Cell Type",
                fill_type = "SampleName", fill_label = "Sample",
                clrs_specific = colors_sample_blood,
                plot_type = "All", details = t_cell_types_str)

# percentage of cell types per dataset
p4 <- plot_pcts(pcts = calc_pcts(data = combined_blood_t[[]],
                                 meta_group_by = "Dataset",
                                 focus_group = "annotated_subclusters"),
                tissue_type = "Blood", plot_value = "Cell Type",
                x_axis = "Dataset", x_axis_label = "Dataset",
                fill_type = "annotated_subclusters", fill_label = "Cell Type",
                clrs_specific = colors_annotated_blood_t,
                details = t_cell_types_str) %>%
        add_info_bar(method = "contains", info_type = "Dataset")

# percentage of datasets per cell type
p5 <- plot_pcts(pcts = calc_pcts(data = combined_blood_t[[]],
                                 meta_group_by = "annotated_subclusters",
                                 focus_group = "Dataset"),
                tissue_type = "Blood", plot_value = "Datasets",
                x_axis = "annotated_subclusters", x_axis_label = "Cell Type",
                fill_type = "Dataset", clrs_specific = colors_dataset,
                details = t_cell_types_str)

# combine all plots
((p1 / p4 + plot_layout(guides = "collect") |
((p2 / p3 / p5) + plot_layout(guides = "collect"))) &
    guides(fill = guide_legend(ncol = 1)))
```

```{r blood-t-check-annotations-cell-type-samples-heatmap}
#| fig.dim = c(8, 8)

groups_cell_type <- levels(combined_blood_t$annotated_subclusters)
groups_sample <- sort(unique(combined_blood_t$SampleName))

overlap_matrix <- matrix(0, nrow = length(groups_sample),
                         ncol = length(groups_cell_type),
                         dimnames = list(groups_sample, groups_cell_type))

for (group_sample in groups_sample) {
  for (group_cell_type in groups_cell_type) {
    cells_cell_type <- combined_blood_t[[]] %>%
                         filter(annotated_subclusters == group_cell_type) %>%
                         pull(Cell_ID_Unique)
    cells_sample <- combined_blood_t[[]] %>%
                      filter(SampleName == group_sample) %>%
                      pull(Cell_ID_Unique)
    
    overlap_count <- length(intersect(cells_cell_type, cells_sample))
    # set zeroes to NAs for plotting purposes
    if (overlap_count == 0) overlap_count <- NA
    
    overlap_matrix[group_sample, group_cell_type] <- overlap_count
  }
}

# to make it easier visually
sample_datasets <- meta %>% select(Dataset, SampleName) %>%
                    filter(SampleName %in% rownames(overlap_matrix)) %>%
                    distinct() %>% pull(Dataset)

ComplexHeatmap::Heatmap(
  matrix = overlap_matrix,
  col = brewer.reds(10)[1:7], # exclude darkest colors
  na_col = "grey75",
  rect_gp = gpar(col = "black"),
  cell_fun = function(j, i, x, y, width, height, fill) {
    if (!is.na(overlap_matrix[i, j])) {
      grid.text(overlap_matrix[i, j], x, y,
                gp = gpar(fontsize = 10, col = "black"))
      }
    },
  row_title = "Sample",
  row_title_side = "right",
  row_title_gp = gpar(fontsize = 12, fontface = "italic"),
  column_title = "Cell Type",
  column_title_side = "bottom",
  column_title_gp = gpar(fontsize = 12, fontface = "italic"),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  row_names_gp = gpar(fontsize = 9),
  column_names_gp = gpar(fontsize = 9),
  left_annotation =
    rowAnnotation(Dataset = sample_datasets,
                      col = list("Dataset" = colors_dataset),
                      annotation_legend_param =
                        list(title_gp = gpar(fontsize = 10),
                             labels_gp = gpar(fontsize = 8)),
                      show_annotation_name = FALSE),
  heatmap_legend_param = list(title = "Overlap",
                              title_gp = gpar(fontsize = 10),
                              labels_gp = gpar(fontsize = 8)),
  use_raster = TRUE) %>%
draw(column_title = "Blood T Cells Cell ID Overlaps by Cell Type and Sample")
```

### Dot plots

Subclustered labels:

```{r blood-t-check-annotations-dot-sub}
#| fig.dim = c(18, 4)

DotPlot(combined_blood_t,
        features = get_features_from_all(markers_df = markers_all,
                                         cell_types = blood_t_marker_types,
                                         alphabetize_all = FALSE),
        cols = colors_dot, dot.scale = 3) %>%
  add_info_bar(method = "join", info_type = "Cell_Type_Full",
               info = get_cell_types(markers_all %>%
                                       filter(Cell_Type %in% blood_t_marker_types)),
               text_size = 6, label = FALSE, angle = 90) %>%
  plot_dot_side(seurat_obj = combined_blood_t,
                row_identity = "annotated_subclusters",
                facet_col = "Cell_Type_Full",
                info_to_add = c("cluster_size", "percent_TCR")) +
  labs(title = "Subclustered Blood T Cells") +
  RotatedAxis() + labels_standard
```

Blood labels:

```{r blood-t-check-annotations-dot}
#| fig.dim = c(18, 6)

Idents(combined_blood_t) <- "annotated_clusters"

DotPlot(combined_blood_t,
        features = get_features_from_all(markers_df = markers_all,
                                         cell_types = blood_t_marker_types,
                                         alphabetize_all = FALSE),
        cols = colors_dot, dot.scale = 3) %>%
  add_info_bar(method = "join", info_type = "Cell_Type_Full",
               info = get_cell_types(markers_all %>%
                                       filter(Cell_Type %in% blood_t_marker_types)),
               text_size = 6, label = FALSE, angle = 90) %>%
  plot_dot_side(seurat_obj = combined_blood_t,
                row_identity = "annotated_clusters",
                facet_col = "Cell_Type_Full",
                info_to_add = c("cluster_size", "percent_TCR")) +
  labs(title = "Subclustered Blood T Cells") +
  RotatedAxis() + labels_standard

Idents(combined_blood_t) <- "annotated_subclusters"
```

## GEX & AIRR

### Cell counts

```{r blood-t-airr-cell-counts-bar}
#| fig.dim = c(12, 8)

ggplot(combined_blood_t[[]] %>%
         select(seurat_subclusters, Has_TCR) %>%
         mutate(Has_TCR = as.character(Has_TCR)),
       aes(x = seurat_subclusters, fill = Has_TCR)) +
  geom_bar(position = "stack", color = "black", linewidth = 0.2) +
  labs(title = "Blood T Cells by Cluster", x = "Cluster", y = "Count") +
  scale_y_continuous(breaks = breaks_pretty()) +
  scale_fill_manual(values = colors_binary) +
  theme_bw + labels_standard
```

### Stacked bar plots {.tabset}

#### TCRs per Sample

```{r blood-t-airr-all-bar}
#| fig.dim = c(12, 8)

plot_pcts(pcts = calc_pcts(data = combined_blood_t[[]],
                           meta_group_by = c("SampleName", "Dataset",
                                             "SampleType"),
                           focus_group = "Has_TCR"),
          tissue_type = "Subclustered Blood", plot_value = "TCRs",
          x_axis = "SampleName", x_axis_label = "Sample",
          fill_type = "Has_TCR",
          clrs_specific = colors_binary,
          plot_type = "Binary", perc_min = 2,
          details = t_cell_types_str) %>%
  add_info_bar(method = "contains", info_type = "Dataset")
```

#### TCRs in T Cells per Sample

```{r blood-t-airr-t-cells-bar}
#| fig.dim = c(12, 8)

plot_pcts(pcts = calc_pcts(data = filter(combined_blood_t[[]],
                                         annotated_clusters %in% t_cell_types),
                           meta_group_by = c("SampleName", "Dataset",
                                             "SampleType"),
                           focus_group = "Has_TCR"),
          tissue_type = "Subclustered Blood", plot_value = "TCRs in T Cells",
          x_axis = "SampleName", x_axis_label = "Sample",
          fill_type = "Has_TCR",
          clrs_specific = colors_binary,
          plot_type = "Binary", perc_min = 2,
          details = t_cell_types_str) %>%
  add_info_bar(method = "contains", info_type = "Dataset")
```

#### TCRs per Cell Type

*Sanity check*

```{r blood-t-airr-cell-types-tcrs-bar}
#| fig.dim = c(12, 8)

plot_pcts(pcts = calc_pcts(data = combined_blood_t[[]],
                           meta_group_by = "annotated_clusters",
                           focus_group = "Has_TCR"),
          tissue_type = "Subclustered Blood", plot_value = "TCRs",
          x_axis = "annotated_clusters", x_axis_label = "Cell Type",
          fill_type = "Has_TCR",
          clrs_specific = colors_binary,
          plot_type = "Binary", perc_min = 2,
          details = t_cell_types_str)
```

#### Cell Types per TCRs

```{r blood-t-airr-tcrs-cell-types-bar}
#| fig.dim = c(12, 8)

plot_pcts(pcts = calc_pcts(data = combined_blood_t[[]],
                           meta_group_by = "Has_TCR",
                           focus_group = "annotated_subclusters"),
          tissue_type = "Subclustered Blood", plot_value = "Cell Types",
          x_axis = "Has_TCR", x_axis_label = "Has TCRs",
          fill_type = "annotated_subclusters", fill_label = "Cell Type",
          clrs_specific = colors_annotated_blood_t,
          perc_min = 2)
```

### UMAPs

#### Has TCRs

All:

```{r blood-t-airr-overlays-tcr-all}
#| fig.dim = c(14, 6)

# plot the UMAP overlays for all samples
p1 <- plot_umap(seurat_obj = combined_blood_t,
                tissue_type = "Blood",
                clrs_specific = colors_annotated_blood_t,
                plot_by = "cluster_all", plot_label = FALSE,
                annotated = TRUE, annotations_col = "annotated_subclusters")
p2 <- plot_immune_overlay(seurat_obj = combined_blood_t, tissue_type = "Blood",
                          airr_type = "TCR", clrs_specific = colors_datatype,
                          plot_by = "all")
p3 <- plot_immune_overlay(seurat_obj = combined_blood_t, tissue_type = "Blood",
                          airr_type = "TCR", clrs_specific = colors_dataset,
                          plot_by = "dataset")

p1 | p2 | p3
```

By Sample:

```{r blood-t-airr-overlays-tcr-sample}
#| fig.dim = c(16, 8)

plot_immune_overlay(seurat_obj = combined_blood_t, tissue_type = "Blood",
                    airr_type = "TCR", clrs_specific = colors_dataset,
                    plot_by = "sample", ncol_sample = 4)
```

#### GEX vs. TCR

```{r blood-t-airr-overlays-gex-tcr}
#| fig.dim = c(14, 6)

p1 <- plot_umap(seurat_obj = combined_blood_t,
                tissue_type = "Blood",
                clrs_specific = colors_annotated_blood,
                plot_by = "cluster_specific",
                specific_clusters =
                  intersect(t_cell_types,
                            levels(combined_blood_t$annotated_clusters)),
                plot_label = FALSE, annotated = TRUE)
p2 <- plot_umap(seurat_obj = combined_blood_t,
                tissue_type = "Subclustered Blood T Cells",
                clrs_specific = colors_annotated_blood_t,
                plot_by = "cluster_all", plot_label = FALSE,
                annotated = TRUE, annotations_col = "annotated_subclusters") +
        labs(color = "")
p3 <- plot_immune_overlay(seurat_obj = combined_blood_t, tissue_type = "Blood",
                          airr_type = "TCR", clrs_specific = colors_datatype,
                          plot_by = "all")

p1 | p2 | p3 & legend_bottom
```

### Venn diagram

With the AIRR data filtered down to just the blood and the β chains:

```{r blood-t-airr-venn}
#| fig.dim = c(9, 7)

ggVennDiagram(
  x = list(GEX = colnames(combined_blood_t),
           TCR = (dplyr::filter(combined_tcr, locus == "TRB",
                                TissueType == "Blood"))$Cell_ID_Unique),
  set_color = colors_datatype[c("GEX", "TCR")]) +
  labs(title = "Subclustered Blood T Cells",
       subtitle = "ID Overlaps across Assays",
       fill = "Cell Count") +
  scale_fill_distiller(palette = "Greys", direction = 1) +
  labels_standard_venn + coord_flip()
```

------------------------------------------------------------------------

# Skin

## Object setup

### Select the clusters

```{r skin-t-setup-subset}
#| eval = FALSE

combined_skin_t <- subset(combined_skin,
                            Has_TCR == TRUE | # AIRR info
                            annotated_clusters %in% t_cell_types) # GEX info
```

```{r skin-t-setup-overview}
subcluster_overview(seurat_obj = combined_skin,
                    seurat_obj_sub = combined_skin_t,
                    tissue_type = "Skin", airr_type = "TCR",
                    cell_types = t_cell_types)
```

### Rerun the standard pipeline

Choose a resolution:

```{r skin-t-seurat-pipeline-res}
skin_t_res <- 0.9
```

With the chosen resolution (with help from `clustree`) of `r skin_t_res`:

```{r skin-t-seurat-pipeline-save}
#| eval = FALSE

# resolution picked from the clustering tree, plus an extra one
combined_skin_t <-
  seurat_pipeline(seurat_obj = combined_skin_t, subcluster = TRUE,
                  num_features = 2000, num_pcs = 30, num_dims = 20, 
                  cluster_res = c(0.6, skin_t_res, 1.1))

# choose a resolution
Idents(combined_skin_t) <- paste0("RNA_snn_sub_res.", skin_t_res)
combined_skin_t$seurat_subclusters <-
  combined_skin_t[[paste0("RNA_snn_sub_res.", skin_t_res)]]

# save the object
saveRDS(combined_skin_t, file.path(path_objs, "combined_skin_t.rds"))
```

### Examine results

```{r skin-t-seurat-pipeline-results-elbow-bar}
#| fig.dim = c(10, 12)

# plot variable features with labels for the top 10
p1 <- LabelPoints(VariableFeaturePlot(combined_skin_t),
                  points = head(VariableFeatures(combined_skin_t), 10),
                  xnudge = 0, ynudge = 0) +
        labels_standard + legend_bottom

p2 <- ElbowPlot(combined_skin_t, ndims = 25) + labels_standard

p3 <- plot_counts_cluster(clusters = combined_skin_t$seurat_subclusters,
                          tissue_type = "Skin",
                          details = "Subclustered T Cells",
                          x_axis = "Cluster")

((p1 | p2) / p3) + plot_layout(heights = c(2, 3))
```

## Dimensionality reduction {.tabset}

### Clusters by Samples

```{r skin-t-clusters-samples-bar}
#| fig.dim = c(16, 15)

# percentage of clusters per sample
p1 <- plot_pcts(pcts = calc_pcts(data = combined_skin_t[[]],
                                 meta_group_by = c("SampleName", "Dataset"),
                                 focus_group = "seurat_subclusters"),
                tissue_type = "Skin", plot_value = "Clusters",
                x_axis = "SampleName", x_axis_label = "Sample",
                fill_type = "seurat_subclusters", fill_label = "Cluster",
                plot_type = "All", details = t_cell_types_str) %>%
        add_info_bar(method = "contains", info_type = "Dataset")

# see which samples are going into each cluster
p2 <- plot_tile(seurat_obj = combined_skin_t, tissue_type = "Skin",
                clrs_specific = colors_sample_skin,
                x_axis = "seurat_subclusters", x_axis_label = "Seurat Cluster",
                details = t_cell_types_str) %>%
        add_info_bar(info_type = "Dataset", top_side = FALSE)

# percentage of samples per cluster
p3 <- plot_pcts(pcts = calc_pcts(data = combined_skin_t[[]],
                                 meta_group_by = "seurat_subclusters",
                                 focus_group = "SampleName"),
                tissue_type = "Skin", plot_value = "Samples",
                x_axis = "seurat_subclusters", x_axis_label = "Cluster",
                fill_type = "SampleName", fill_label = "Sample",
                clrs_specific = colors_sample_skin,
                plot_type = "All", details = t_cell_types_str)

# percentage of clusters per dataset
p4 <- plot_pcts(pcts = calc_pcts(data = combined_skin_t[[]],
                                 meta_group_by = "Dataset",
                                 focus_group = "seurat_subclusters"),
                tissue_type = "Skin", plot_value = "Clusters",
                x_axis = "Dataset", x_axis_label = "Dataset",
                fill_type = "seurat_subclusters", fill_label = "Cluster",
                details = t_cell_types_str) %>%
        add_info_bar(method = "contains", info_type = "Dataset")

# percentage of datasets per cluster
p5 <- plot_pcts(pcts = calc_pcts(data = combined_skin_t[[]],
                                 meta_group_by = "seurat_subclusters",
                                 focus_group = "Dataset"),
                tissue_type = "Skin", plot_value = "Datasets",
                x_axis = "seurat_subclusters", x_axis_label = "Cluster",
                fill_type = "Dataset", clrs_specific = colors_dataset,
                details = t_cell_types_str)

# combine all plots
((p1 / p4 + plot_layout(guides = "collect") |
((p2 / p3 / p5) + plot_layout(guides = "collect"))) &
    guides(fill = guide_legend(ncol = 1)))
```

### Cells by Samples

```{r skin-t-cells-samples-bar}
#| fig.dim = c(12, 8)

plot_counts_cluster(clusters = combined_skin_t$SampleName,
                    tissue_type = "Skin",
                    details = "Subclustered T Cells",
                    x_axis = "Cell Type",
                    clrs_specific = colors_samples) +
  labels_rotate_x
```

### Cell Types by Samples

```{r skin-t-cell-types-samples-bar}
#| fig.dim = c(12, 8)

plot_counts_cluster(clusters = combined_skin_t$annotated_clusters,
                    tissue_type = "Skin",
                    details = "Subclustered T Cells",
                    x_axis = "Cell Type",
                    clrs_specific = colors_annotated_skin) +
  labels_rotate_x
```

### Samples

```{r skin-t-samples-bar}
#| fig.dim = c(12, 7)

plot_pcts(calc_pcts(data = combined_skin_t[[]] %>% mutate(Type = "Sample"),
                    meta_group_by = "Type",
                    focus_group = "SampleName"),
          tissue_type = "Subclustered Skin T Cells",
          clrs_specific = colors_sample_skin,
          plot_value = "Sample", x_axis = "Type", x_axis_label = "Total",
          fill_type = "SampleName", fill_label = "Sample",
          perc_min = 2, label_size = 3, label_fill = TRUE,
          include_counts = FALSE) +
  coord_flip() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(ncol = 1))
```

### UMAPs

All:

```{r skin-t-umaps}
#| fig.dim = c(24, 8)

p1 <- plot_umap(seurat_obj = combined_skin_t,
                tissue_type = "Subclustered Skin T Cells",
                use_hues = TRUE, plot_by = "cluster_all",
                clusters_col = "seurat_subclusters")
p2 <- plot_umap(seurat_obj = combined_skin_t,
                tissue_type = "Subclustered Skin T Cells",
                clrs_specific = colors_sample_skin,
                plot_by = "all", plot_label = FALSE,
                clusters_col = "seurat_subclusters")
p3 <- plot_umap(seurat_obj = combined_skin_t,
                tissue_type = "Subclustered Skin T Cells",
                clrs_specific = colors_dataset,
                plot_by = "dataset", plot_label = FALSE,
                clusters_col = "seurat_subclusters")

p1 | p2 | p3
```

By Sample:

```{r skin-t-umap-cluster-sample}
#| fig.dim = c(12, 18)

plot_umap(seurat_obj = combined_skin_t,
          tissue_type = "Subclustered Skin T Cells",
          use_hues = TRUE, plot_by = "cluster_sample",
          clusters_col = "seurat_subclusters")
```

## Marker overlays

### Setup

```{r skin-t-marker-overlays-setup}
#| class.source = "fold-show"

# all T cell markers ("T" is too broad)
skin_t_marker_types <- c("NK", "T", "T_CD4", "T_CD4RM", "T_CD8", "T_CD8RM",
                         "T_Dividing", "T_FH", "T_Reg", "TCM", "TCR",
                         "TEM", "TRM", "TteK")

get_features_from_all(markers_df = markers_all, cell_types = skin_t_marker_types)
```

### Dot plot

```{r skin-t-marker-overlays-dot}
#| fig.dim = c(20, 6)

DotPlot(combined_skin_t,
        features = get_features_from_all(markers_df = markers_all,
                                         cell_types = skin_t_marker_types,
                                         alphabetize_all = FALSE),
        cols = colors_dot, dot.scale = 3, cluster.idents = TRUE) %>%
  add_info_bar(method = "join", info_type = "Cell_Type_Full",
               info = get_cell_types(markers_all %>%
                                       filter(Cell_Type %in% skin_t_marker_types)),
               text_size = 6, label = FALSE, angle = 90) %>%
  plot_dot_side(seurat_obj = combined_skin_t,
                row_identity = "seurat_subclusters",
                facet_col = "Cell_Type_Full",
                info_to_add = c("cluster_size", "percent_TCR")) +
  labs(title = "Subclustered Skin T Cells") +
  RotatedAxis() + labels_standard
```

### Feature plots

```{r skin-t-marker-overlays-feature}
#| fig.dim = c(16, 52)

# this spits out a lot of plots
FeaturePlot(combined_skin_t,
            features = get_features_from_all(markers_df = markers_all,
                                             cell_types = skin_t_marker_types),
            min.cutoff = 0,
            label = TRUE, label.size = 2, ncol = 4, raster = FALSE) +
  plot_annotation(title = "Subclustered Skin T Cells") &
  labels_standard & clean_umap
```

## Annotate cell clusters

### Manually

```{r skin-t-annotations-manual}
#| eval = FALSE

combined_skin_t_annotations <- rbind(c("0", "CD8+ T Cells"),
                                     c("1", "TBD T Cells"),
                                     c("2", "TBD T Cells"),
                                     c("3", "CD4+ T Cells"),
                                     c("4", "Regulatory T Cells"), # FOXP3, CD25/IL2RA
                                     c("5", "CD8+ T Cells"), # CD8
                                     c("6", "Cytotoxic T Cells"), # NK T cells?
                                     c("7", "CD4+ T Cells"), # CCR7
                                     c("8", "TBD T Cells"), # CD3E
                                     c("9", "TBD T Cells"),
                                     c("10", "TBD T Cells"),
                                     c("11", "Dividing T Cells"), # MKI67
                                     c("12", "Non T Cells"), # monocytes
                                     c("13", "CD8+ T Cells"),
                                     c("14", "Non T Cells"), # fibroblasts
                                     c("15", "Non T Cells"), # keratinocytes
                                     c("16", "Non T Cells"), # B cells
                                     c("17", "Non T Cells")) # endothelial
colnames(combined_skin_t_annotations) <- c("Cluster", "CellType")
combined_skin_t_annotations <- data.frame(combined_skin_t_annotations)

# save the annotations as a csv
write_csv(combined_skin_t_annotations,
          file.path(path_tables, "annotations_skin_t.csv"))
```

Add the annotations to the Seurat object and examine them:

```{r skin-t-annotations-manual-add}
# read in and show the annotations
annotations_skin_t <- read_csv(file.path(path_tables,
                                         "annotations_skin_t.csv"),
                               col_types = "c")

# add the annotations to the object
combined_skin_t <- add_annotations(seurat_obj = combined_skin_t,
                                   annotations_df = annotations_skin_t,
                                   clusters_col = "seurat_subclusters",
                                   annotations_col = "annotated_subclusters")

# clusters by annotation
print_kable(annotations_skin_t %>%
              group_by(CellType) %>%
              transmute(Clusters = paste0(Cluster, collapse = ", ")) %>%
              distinct() %>% arrange(CellType),
            kable_height = NULL)
```

Choose the color scheme:

```{r skin-t-annotations-color-schemes}
# only want the non B cells to be black
skin_t_cell_types_num <- n_distinct(combined_skin_t$annotated_subclusters)
skin_t_cell_types <- sort(unique(combined_skin_t$annotated_subclusters))

colors_annotated_skin_t <-
    setNames(rep("#000000", skin_t_cell_types_num),
             nm = skin_t_cell_types)
colors_annotated_skin_t[str_subset(skin_t_cell_types, "Non", negate = TRUE)] <-
  iwanthue(n = skin_t_cell_types_num - 1, hmin = 60, hmax = 190)

# save the colors to the miscellaneous slot for easy access
Misc(combined_skin_t, slot = "colors_annotated") <- colors_annotated_skin_t
```

Save the annotated object:

```{r skin-t-annotations-manual-save}
#| eval = FALSE

# remove unneeded columns from the metadata
for (col in c("sequence_id_unique", "v_call_family_TRA", "j_call_family_TRA")) {
  combined_skin_t[[col]] <- c()
}

# save the object with annotations (watch out for reproducibility)
saveRDS(combined_skin_t, file.path(path_objs, "combined_skin_t.rds"))
```

### FindMarkers

All clusters:

```{r skin-t-annotations-findallmarkers-save}
#| eval = FALSE

Idents(combined_skin_t) <- "seurat_subclusters"

# find markers for every cluster compared to all remaining cells
# report only the positive ones
combined_skin_t_findallmarkers <- FindAllMarkers(combined_skin_t,
                                                 only.pos = TRUE, min.pct = 0.25,
                                                 logfc.threshold = 0.25)

Idents(combined_skin_t) <- "annotated_subclusters"

# save the markers as a csv
write_csv(combined_skin_t_findallmarkers,
          file.path(path_tables, "findmarkers",
                    "annotations_skin_t_findallmarkers.csv"))
```

```{r skin-t-annotations-findallmarkers-load}
# load markers
combined_skin_t_findallmarkers <-
  read_csv(file.path(path_tables, "findmarkers",
                     "annotations_skin_t_findallmarkers.csv"),
           show_col_types = FALSE)

print_kable(combined_skin_t_findallmarkers %>%
              group_by(cluster) %>%
              slice_max(n = 5, order_by = desc(p_val_adj)) %>%
              relocate(c("cluster", "gene", "p_val_adj"), .before = c("p_val")))
```

### Final annotations

```{r skin-t-annotations-manual-cluster-all-umap}
#| fig.dim = c(24, 8)

p1 <- plot_umap(seurat_obj = combined_skin_t,
                tissue_type = "Subclustered Skin T Cells",
                plot_by = "cluster_all",
                clusters_col = "seurat_subclusters") # for comparison
p2 <- plot_umap(seurat_obj = combined_skin_t,
                tissue_type = "Subclustered Skin T Cells",
                clrs_specific = colors_annotated_skin,
                plot_by = "cluster_all",
                annotated = TRUE, annotations_col = "annotated_clusters",
                details = "Skin Cell Type")
p3 <- plot_umap(seurat_obj = combined_skin_t,
                tissue_type = "Subclustered Skin T Cells",
                clrs_specific = colors_annotated_skin_t,
                plot_by = "cluster_all",
                annotated = TRUE, annotations_col = "annotated_subclusters")

p1 | p2 | p3
```

```{r skin-t-annotations-manual-cluster-sample-umap}
#| fig.dim = c(12, 18)

plot_umap(seurat_obj = combined_skin_t,
          tissue_type = "Subclustered Skin T Cells",
          clrs_specific = colors_annotated_skin_t,
          plot_by = "cluster_sample",
          annotated = TRUE, annotations_col = "annotated_subclusters")
```

## Check the annotations

### Cell counts per cluster

```{r skin-t-check-annotations-cell-type-counts-bar}
#| fig.dim = c(10, 6)

plot_counts_cluster(clusters = combined_skin_t$annotated_subclusters,
                    tissue_type = "Skin", details = "Subclustered T Cells",
                    x_axis = "Cell Type",
                    clrs_specific = colors_annotated_skin_t)
```

```{r skin-t-check-annotations-cell-type-counts-stacked-bar}
#| fig.dim = c(15, 5)

plot_pcts(calc_pcts(data = combined_skin_t[[]] %>% mutate(Type = "Sample"),
                    meta_group_by = "Type",
                    focus_group = "annotated_subclusters"),
          tissue_type = "Skin T Cells",
          clrs_specific = colors_annotated_skin_t,
          plot_value = "Cell Types", x_axis = "Type", x_axis_label = "Total",
          fill_type = "annotated_subclusters", fill_label = "Cell Type",
          perc_min = 1.5, label_size = 3, label_fill = TRUE,
          include_counts = FALSE, reverse_order = TRUE) +
  coord_flip() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(ncol = 1))
```

### Cell Types by Samples

```{r skin-t-check-annotations-cell-types-samples-bar}
#| fig.dim = c(16, 15)

# percentage of cell types per sample
p1 <- plot_pcts(pcts = calc_pcts(data = combined_skin_t[[]],
                                 meta_group_by = c("SampleName", "Dataset"),
                                 focus_group = "annotated_subclusters"),
                tissue_type = "Skin", plot_value = "Cell Types",
                x_axis = "SampleName", x_axis_label = "Sample",
                fill_type = "annotated_subclusters", fill_label = "Cell Type",
                clrs_specific = colors_annotated_skin_t,
                plot_type = "All", details = t_cell_types_str) %>%
        add_info_bar(method = "contains", info_type = "Dataset")

# see which samples are going into each cell type
p2 <- plot_tile(seurat_obj = combined_skin_t, tissue_type = "Skin",
                clrs_specific = colors_sample_skin,
                x_axis = "annotated_subclusters", x_axis_label = "Cell Type",
                details = t_cell_types_str) %>%
        add_info_bar(info_type = "Dataset", top_side = FALSE)

# percentage of samples per cell type
p3 <- plot_pcts(pcts = calc_pcts(data = combined_skin_t[[]],
                                 meta_group_by = "annotated_subclusters",
                                 focus_group = "SampleName"),
                tissue_type = "Skin", plot_value = "Samples",
                x_axis = "annotated_subclusters", x_axis_label = "Cell Type",
                fill_type = "SampleName", fill_label = "Sample",
                clrs_specific = colors_sample_skin,
                plot_type = "All", details = t_cell_types_str)

# percentage of cell types per dataset
p4 <- plot_pcts(pcts = calc_pcts(data = combined_skin_t[[]],
                                 meta_group_by = "Dataset",
                                 focus_group = "annotated_subclusters"),
                tissue_type = "Skin", plot_value = "Cell Types",
                x_axis = "Dataset", x_axis_label = "Dataset",
                fill_type = "annotated_subclusters", fill_label = "Cell Type",
                clrs_specific = colors_annotated_skin_t,
                details = t_cell_types_str) %>%
        add_info_bar(method = "contains", info_type = "Dataset")

# percentage of datasets per cell type
p5 <- plot_pcts(pcts = calc_pcts(data = combined_skin_t[[]],
                                 meta_group_by = "annotated_subclusters",
                                 focus_group = "Dataset"),
                tissue_type = "Skin", plot_value = "Datasets",
                x_axis = "annotated_subclusters", x_axis_label = "Cell Type",
                fill_type = "Dataset", clrs_specific = colors_dataset,
                details = t_cell_types_str)

# combine all plots
((p1 / p4 + plot_layout(guides = "collect") |
((p2 / p3 / p5) + plot_layout(guides = "collect"))) &
    guides(fill = guide_legend(ncol = 1)))
```

```{r skin-t-check-annotations-cell-type-samples-heatmap}
#| fig.dim = c(8, 8)

groups_cell_type <- levels(combined_skin_t$annotated_subclusters)
groups_sample <- sort(unique(combined_skin_t$SampleName))

overlap_matrix <- matrix(0, nrow = length(groups_sample),
                         ncol = length(groups_cell_type),
                         dimnames = list(groups_sample, groups_cell_type))

for (group_sample in groups_sample) {
  for (group_cell_type in groups_cell_type) {
    cells_cell_type <- combined_skin_t[[]] %>%
                         filter(annotated_subclusters == group_cell_type) %>%
                         pull(Cell_ID_Unique)
    cells_sample <- combined_skin_t[[]] %>%
                      filter(SampleName == group_sample) %>%
                      pull(Cell_ID_Unique)
    
    overlap_count <- length(intersect(cells_cell_type, cells_sample))
    # set zeroes to NAs for plotting purposes
    if (overlap_count == 0) overlap_count <- NA
    
    overlap_matrix[group_sample, group_cell_type] <- overlap_count
  }
}

# to make it easier visually
sample_datasets <- meta %>% select(Dataset, SampleName) %>%
                    filter(SampleName %in% rownames(overlap_matrix)) %>%
                    distinct() %>% pull(Dataset)

ComplexHeatmap::Heatmap(
  matrix = overlap_matrix,
  col = brewer.orrd(10)[1:7], # exclude darkest colors
  na_col = "grey75",
  rect_gp = gpar(col = "black"),
  cell_fun = function(j, i, x, y, width, height, fill) {
    if (!is.na(overlap_matrix[i, j])) {
      grid.text(overlap_matrix[i, j], x, y,
                gp = gpar(fontsize = 10, col = "black"))
      }
    },
  row_title = "Sample",
  row_title_side = "right",
  row_title_gp = gpar(fontsize = 12, fontface = "italic"),
  column_title = "Cell Type",
  column_title_side = "bottom",
  column_title_gp = gpar(fontsize = 12, fontface = "italic"),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  row_names_gp = gpar(fontsize = 9),
  column_names_gp = gpar(fontsize = 9),
  left_annotation =
    rowAnnotation(Dataset = sample_datasets,
                      col = list("Dataset" = colors_dataset),
                      annotation_legend_param =
                        list(title_gp = gpar(fontsize = 10),
                             labels_gp = gpar(fontsize = 8)),
                      show_annotation_name = FALSE),
  heatmap_legend_param = list(title = "Overlap",
                              title_gp = gpar(fontsize = 10),
                              labels_gp = gpar(fontsize = 8)),
  use_raster = TRUE) %>%
draw(column_title = "Skin T Cells Cell ID Overlaps by Cell Type and Sample")
```

### Dot plots

Subclustered labels:

```{r skin-t-check-annotations-dot-sub}
#| fig.dim = c(20, 4)

DotPlot(combined_skin_t,
        features = get_features_from_all(markers_df = markers_all,
                                         cell_types = skin_t_marker_types,
                                         alphabetize_all = FALSE),
        cols = colors_dot, dot.scale = 3) %>%
  add_info_bar(method = "join", info_type = "Cell_Type_Full",
               info = get_cell_types(markers_all %>%
                                       filter(Cell_Type %in% skin_t_marker_types)),
               text_size = 6, label = FALSE, angle = 90) %>%
  plot_dot_side(seurat_obj = combined_skin_t,
                row_identity = "annotated_subclusters",
                facet_col = "Cell_Type_Full",
                info_to_add = c("cluster_size", "percent_TCR")) +
  labs(title = "Subclustered Skin T Cells") +
  RotatedAxis() + labels_standard
```

Skin labels:

```{r skin-t-check-annotations-dot}
#| fig.dim = c(20, 6)

Idents(combined_skin_t) <- "annotated_clusters"

DotPlot(combined_skin_t,
        features = get_features_from_all(markers_df = markers_all,
                                         cell_types = skin_t_marker_types,
                                         alphabetize_all = FALSE),
        cols = colors_dot, dot.scale = 3) %>%
  add_info_bar(method = "join", info_type = "Cell_Type_Full",
               info = get_cell_types(markers_all %>%
                                       filter(Cell_Type %in% skin_t_marker_types)),
               text_size = 6, label = FALSE, angle = 90) %>%
  plot_dot_side(seurat_obj = combined_skin_t,
                row_identity = "annotated_clusters",
                facet_col = "Cell_Type_Full",
                info_to_add = c("cluster_size", "percent_TCR")) +
  labs(title = "Subclustered Skin T Cells") +
  RotatedAxis() + labels_standard

Idents(combined_skin_t) <- "annotated_subclusters"
```

## GEX & AIRR

### Cell counts

```{r skin-t-airr-cell-counts-bar}
#| fig.dim = c(12, 8)

ggplot(combined_skin_t[[]] %>%
         select(seurat_subclusters, Has_TCR) %>%
         mutate(Has_TCR = as.character(Has_TCR)),
       aes(x = seurat_subclusters, fill = Has_TCR)) +
  geom_bar(position = "stack", color = "black", linewidth = 0.2) +
  labs(title = "Skin T Cells by Cluster", x = "Cluster", y = "Count") +
  scale_y_continuous(breaks = breaks_pretty()) +
  scale_fill_manual(values = colors_binary) +
  theme_bw + labels_standard
```

### Stacked bar plots {.tabset}

#### TCRs per Sample

```{r skin-t-airr-all-bar}
#| fig.dim = c(12, 8)

plot_pcts(pcts = calc_pcts(data = combined_skin_t[[]],
                           meta_group_by = c("SampleName", "Dataset",
                                             "SampleType"),
                           focus_group = "Has_TCR"),
          tissue_type = "Subclustered Skin", plot_value = "TCRs",
          x_axis = "SampleName", x_axis_label = "Sample",
          fill_type = "Has_TCR",
          clrs_specific = colors_binary,
          plot_type = "Binary", perc_min = 2,
          details = t_cell_types_str) %>%
  add_info_bar(method = "contains", info_type = "Dataset")
```

#### TCRs in T Cells per Sample

```{r skin-t-airr-t-cells-bar}
#| fig.dim = c(12, 8)

plot_pcts(pcts = calc_pcts(data = filter(combined_skin_t[[]],
                                         annotated_clusters %in% t_cell_types),
                           meta_group_by = c("SampleName", "Dataset",
                                             "SampleType"),
                           focus_group = "Has_TCR"),
          tissue_type = "Subclustered Skin", plot_value = "TCRs in T Cells",
          x_axis = "SampleName", x_axis_label = "Sample",
          fill_type = "Has_TCR",
          clrs_specific = colors_binary,
          plot_type = "Binary", perc_min = 2,
          details = t_cell_types_str) %>%
  add_info_bar(method = "contains", info_type = "Dataset")
```

#### TCRs per Cell Type

*Sanity check*

```{r skin-t-airr-cell-types-tcrs-bar}
#| fig.dim = c(12, 8)

plot_pcts(pcts = calc_pcts(data = combined_skin_t[[]],
                           meta_group_by = "annotated_clusters",
                           focus_group = "Has_TCR"),
          tissue_type = "Subclustered Skin", plot_value = "TCRs",
          x_axis = "annotated_clusters", x_axis_label = "Cell Type",
          fill_type = "Has_TCR",
          clrs_specific = colors_binary,
          plot_type = "Binary", perc_min = 2,
          details = t_cell_types_str)
```

#### Cell Types per TCRs

```{r skin-t-airr-tcrs-cell-types-bar}
#| fig.dim = c(12, 8)

plot_pcts(pcts = calc_pcts(data = combined_skin_t[[]],
                           meta_group_by = "Has_TCR",
                           focus_group = "annotated_subclusters"),
          tissue_type = "Subclustered Skin", plot_value = "Cell Types",
          x_axis = "Has_TCR", x_axis_label = "Has TCRs",
          fill_type = "annotated_subclusters", fill_label = "Cell Type",
          clrs_specific = colors_annotated_skin_t,
          perc_min = 2)
```

### UMAPs

#### Has TCRs

All:

```{r skin-t-airr-overlays-tcr-all}
#| fig.dim = c(14, 6)

# plot the UMAP overlays for all samples
p1 <- plot_umap(seurat_obj = combined_skin_t,
                tissue_type = "Skin", clrs_specific = colors_annotated_skin_t,
                plot_by = "cluster_all",
                annotated = TRUE, annotations_col = "annotated_subclusters")
p2 <- plot_immune_overlay(seurat_obj = combined_skin_t, tissue_type = "Skin",
                          airr_type = "TCR", clrs_specific = colors_datatype,
                          plot_by = "all")
p3 <- plot_immune_overlay(seurat_obj = combined_skin_t, tissue_type = "Skin",
                          airr_type = "TCR", clrs_specific = colors_dataset,
                          plot_by = "dataset")

p1 | p2 | p3
```

By Sample:

```{r skin-t-airr-overlays-tcr-sample}
#| fig.dim = c(16, 24)

plot_immune_overlay(seurat_obj = combined_skin_t, tissue_type = "Skin",
                    airr_type = "TCR", clrs_specific = colors_dataset,
                    plot_by = "sample", ncol_sample = 4)
```

#### GEX vs. TCR

```{r skin-t-airr-overlays-gex-tcr}
#| fig.dim = c(14, 6)

p1 <- plot_umap(seurat_obj = combined_skin_t,
                tissue_type = "Skin",
                clrs_specific = colors_annotated_skin,
                plot_by = "cluster_specific", plot_label = FALSE,
                specific_clusters =
                  intersect(t_cell_types,
                            levels(combined_skin_t$annotated_clusters)),
                annotated = TRUE)
p2 <- plot_umap(seurat_obj = combined_skin_t,
                tissue_type = "Subclustered Skin T Cells",
                clrs_specific = colors_annotated_skin_t,
                plot_by = "cluster_all", plot_label = FALSE,
                annotated = TRUE, annotations_col = "annotated_subclusters") +
        labs(color = "")
p3 <- plot_immune_overlay(seurat_obj = combined_skin_t, tissue_type = "Skin",
                          airr_type = "TCR", clrs_specific = colors_datatype,
                          plot_by = "all")
p1 | p2 | p3 & legend_bottom
```

### Venn diagram

With the AIRR data filtered down to just the skin and the β chains:

```{r skin-t-airr-venn}
#| fig.dim = c(9, 7)

ggVennDiagram(
  x = list(GEX = colnames(combined_skin_t),
           TCR = (dplyr::filter(combined_tcr, locus == "TRB",
                                TissueType == "Skin"))$Cell_ID_Unique),
  set_color = colors_datatype[c("GEX", "TCR")]) +
  labs(title = "Subclustered Skin T Cells",
       subtitle = "ID Overlaps across Assays",
       fill = "Cell Count") +
  scale_fill_distiller(palette = "Greys", direction = 1) +
  labels_standard_venn + coord_flip()
```
