---
title: "Pathway Analysis"
output:
  html_document:
    code_download: yes
    code_folding: hide
    css: ../../"style.css"
    dev: "png"
    df_print: kable
    fig_caption: yes
    keep_md: no
    self_contained: true
    theme: yeti
    toc: yes
    toc_depth: 4
    toc_float:
        collapsed: false
        smooth_scroll: false
editor_options:
  chunk_output_type: console
---

------------------------------------------------------------------------

# Overview and setup

```{r setup-options, include = FALSE}
knitr::opts_chunk$set(comment = NA, dpi = 200, fig.align = "center",
                      out.width = "100%", warning = FALSE, error = TRUE,
                      message = FALSE,
                      fig.path = here::here("figures",
                                            "downstream", "pathway_analysis",
                                            "pathways-"))

# will mess with chunk labels, but needed for parameterized reports
options(knitr.duplicate.label = "allow")
```

**Author(s):** Edel Aron<br>
**Last Updated:** `r Sys.Date()`

**R version:** `r R.Version()$version.string`<br>
**Platform:** `r R.Version()$platform`<br>
**Running under:** `r sessionInfo()$running`

```{r, include = FALSE}
# load packages, colors and themes, functions, and metadata (R code only)
code_setup <-
  knitr::purl(here::here("notebooks", "01-setup.Rmd"),
              output = tempfile(), documentation = 0)
code_functions <-
  knitr::purl(here::here("notebooks", "02-functions.Rmd"),
              output = tempfile(), documentation = 0)
code_meta <-
  knitr::purl(here::here("notebooks", "03-meta.Rmd"),
              output = tempfile(), documentation = 0)

# read in the code, then remove the temp files
source(code_setup)
source(code_functions)
source(code_meta)
unlink(c(code_setup, code_functions, code_meta))
```

Connect to `Enrichr`:

```{r setup-enrichr}
setEnrichrSite("Enrichr")
```

Additional libraries:

```{r}
packages <- c()
for (n in seq_along(packages)) {
  suppressPackageStartupMessages(library(packages[n], character.only = TRUE))
  cat(paste0(packages[n], ": ", packageVersion(packages[n]), "\n"))
}
```

Could also use tools such as clusterProfiler, msigdbr, or ggpubr (for volcano plots).

------------------------------------------------------------------------

# Read in the data

## GEX

```{r setup}
# objects
combined_skin <-
  readRDS(file.path(path_objs, "final", "combined_skin_airr.rds"))
combined_skin_recombined_t <-
  readRDS(file.path(path_objs, "final", "combined_skin_recombined_t.rds"))

# colors
colors_annotated_skin <- combined_skin@misc$colors_annotated
colors_annotated_skin_recombined_t <- combined_skin_recombined_t@misc$colors_annotated
```

Subclustered objects use the refined cells:

```{r setup-subcluster-skin}
# objects
combined_skin_t <-
  readRDS(file.path(path_objs, "final", "combined_skin_t_ref.rds"))

# colors
colors_annotated_skin_t <- combined_skin_t@misc$colors_annotated
```

## Differential expression

Define the FDR cutoff:

```{r setup-de-setup}
#| class.source = "fold-show"

# use the same one for everything
fdr_cutoff <- 0.05
```

Read in the DESeq2 results:

```{r setup-de-deseq2}
# by cell type
de_skin <-
  readRDS(file.path(path_objs, "downstream", "differential_expression",
                    "deseq2_combined_skin.rds"))
de_skin_recombined_t <-
  readRDS(file.path(path_objs, "downstream", "differential_expression",
                    "deseq2_combined_skin_recombined_t.rds"))
de_skin_t <-
  readRDS(file.path(path_objs, "downstream", "differential_expression",
                    "deseq2_combined_skin_t.rds"))
```

Create the filtered tables:

```{r setup-de-tables}
de_genes_skin <- enframe(de_skin) %>%
                   unnest(value) %>%
                   rename(cell_type = name) %>%
                   dplyr::filter(p_val_adj < fdr_cutoff)
de_genes_skin_recombined_t <- enframe(de_skin_recombined_t) %>%
                              unnest(value) %>%
                              rename(cell_type = name) %>%
                              dplyr::filter(p_val_adj < fdr_cutoff)
de_genes_skin_t <- enframe(de_skin_t) %>%
                     unnest(value) %>%
                     rename(cell_type = name) %>%
                     dplyr::filter(p_val_adj < fdr_cutoff)
```

------------------------------------------------------------------------

# Differential expression {.tabset}

Set up the required columns needed to calculate pseudobulk and run DE:

```{r setup-de-setup-cols}
combined_skin$replicate <- combined_skin$Subject
combined_skin$label <- combined_skin$SampleType
combined_skin$cell_type <- combined_skin$annotated_clusters

combined_skin_recombined_t$replicate <- combined_skin_recombined_t$Subject
combined_skin_recombined_t$label <- combined_skin_recombined_t$SampleType
combined_skin_recombined_t$cell_type <-
  combined_skin_recombined_t$annotated_clusters_expanded

combined_skin_t$replicate <- combined_skin_t$Subject
combined_skin_t$label <- combined_skin_t$SampleType
combined_skin_t$cell_type <- combined_skin_t$annotated_subclusters
```

Running differential expression on all of the cells together is a little nonsensical, so we run it on each cell type instead.

[Recommendations for single-cell analysis](https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#recommendations-for-single-cell-analysis)

## Skin

```{r de-skin-calc}
#| eval = FALSE

de_skin <- pseudobulk_DE(combined_skin, de_test = "pseudobulk_DESeq2")

saveRDS(de_skin,
        file.path(path_objs, "downstream", "differential_expression",
                  "deseq2_combined_skin.rds"))
```

## Skin (Recombined T Cells)

```{r de-skin-recombined-t-calc}
#| eval = FALSE

de_skin_recombined_t <-
  pseudobulk_DE(combined_skin_recombined_t, de_test = "pseudobulk_DESeq2")

saveRDS(de_skin_recombined_t,
        file.path(path_objs, "downstream", "differential_expression",
                  "deseq2_combined_skin_recombined_t.rds"))
```

## Skin T Cells

```{r de-skin-t-calc}
#| eval = FALSE

de_skin_t <- pseudobulk_DE(combined_skin_t, de_test = "pseudobulk_DESeq2")

saveRDS(de_skin_t,
        file.path(path_objs, "downstream", "differential_expression",
                  "deseq2_combined_skin_t.rds"))
```

## Skin T Cells (EM only)

CD8+ GZMK+ IFNG exploration:

```{r de-skin-t-em-ifng-calc}
#| eval = FALSE

combined_skin_t_em_ifng <-
  subset(combined_skin_t, subset = SampleType == "EM")
combined_skin_t_em_ifng <-
  subset(combined_skin_t_em_ifng,
         subset = annotated_subclusters %in% c("CD8+ GZMK+ IFNGhi T Cells",
                                               "CD8+ GZMK+ IFNGint T Cells"))
combined_skin_t_em_ifng$annotated_subclusters <-
  droplevels(combined_skin_t_em_ifng$annotated_subclusters)

combined_skin_t_em_ifng$replicate <- combined_skin_t_em_ifng$Subject # SampleName
combined_skin_t_em_ifng$label <- combined_skin_t_em_ifng$annotated_subclusters
combined_skin_t_em_ifng$cell_type <- "CD8+ GZMK+ IFNG T Cells"

# we want the IFNGint to be the "reference" cell type
combined_skin_t_em_ifng$label <-
  relevel(combined_skin_t_em_ifng$label, ref = "CD8+ GZMK+ IFNGint T Cells")

de_skin_t_em_ifng <-
  pseudobulk_DE(combined_skin_t_em_ifng, de_test = "pseudobulk_DESeq2")

saveRDS(de_skin_t_em_ifng,
        file.path(path_objs, "downstream", "differential_expression",
                  "deseq2_combined_skin_t_em_ifng.rds"))
```

CD8+ GZMK+ IFNGhi vs. HSPhi T cells:

```{r de-skin-t-em-ifnghi-hsphi-calc}
#| eval = FALSE

combined_skin_t_em_ifnghi_hsphi <-
  subset(combined_skin_t, subset = SampleType == "EM")
combined_skin_t_em_ifnghi_hsphi <-
  subset(combined_skin_t_em_ifnghi_hsphi,
         subset = annotated_subclusters %in% c("CD8+ GZMK+ IFNGhi T Cells",
                                               "HSPhi T Cells"))
combined_skin_t_em_ifnghi_hsphi$annotated_subclusters <-
  droplevels(combined_skin_t_em_ifnghi_hsphi$annotated_subclusters)

combined_skin_t_em_ifnghi_hsphi$replicate <-
  combined_skin_t_em_ifnghi_hsphi$Subject
combined_skin_t_em_ifnghi_hsphi$label <-
  combined_skin_t_em_ifnghi_hsphi$annotated_subclusters
combined_skin_t_em_ifnghi_hsphi$cell_type <- "T Cells"

# we want the HSPhi to be the "reference" cell type
combined_skin_t_em_ifnghi_hsphi$label <-
  relevel(combined_skin_t_em_ifnghi_hsphi$label, ref = "HSPhi T Cells")

de_skin_t_em_ifnghi_hsphi <-
  pseudobulk_DE(combined_skin_t_em_ifnghi_hsphi, de_test = "pseudobulk_DESeq2")

saveRDS(de_skin_t_em_ifnghi_hsphi,
        file.path(path_objs, "downstream", "differential_expression",
                  "deseq2_combined_skin_t_em_ifnghi_hsphi.rds"))
```

Expanded vs. singletons:

Like in [Single-cell T cell receptor sequencing of paired human atherosclerotic plaques and blood reveals autoimmune-like features of expanded effector T cells](https://www.nature.com/articles/s44161-022-00208-4#Fig3):

```{r}
# let's do EM only
combined_skin_t_em_ifnghi <-
  subset(combined_skin_t, subset = SampleType == "EM")
combined_skin_t_em_ifnghi <-
  subset(combined_skin_t_em_ifnghi,
         subset = annotated_subclusters == "CD8+ GZMK+ IFNGhi T Cells")
combined_skin_t_em_ifnghi$annotated_subclusters <-
  droplevels(combined_skin_t_em_ifnghi$annotated_subclusters)

# singletons vs expanded clones
clone_info <- countClones(combined_skin_t_em_ifnghi[[]],
                          clone = "clone_id_unique", remove_na = TRUE) %>%
                mutate(clone_type = ifelse(seq_count > 1, "Expanded", "Singleton"))
combined_skin_t_em_ifnghi@meta.data <-
  left_join(combined_skin_t_em_ifnghi[[]],
            clone_info %>% select(clone_id_unique, clone_type),
            by = join_by(clone_id_unique))
rownames(combined_skin_t_em_ifnghi@meta.data) <-
  combined_skin_t_em_ifnghi$Cell_ID_Unique
combined_skin_t_em_ifnghi <-
  subset(combined_skin_t_em_ifnghi, clone_type %in% c("Expanded", "Singleton"))
combined_skin_t_em_ifnghi$clone_type <-
  factor(combined_skin_t_em_ifnghi$clone_type)

combined_skin_t_em_ifnghi$replicate <-
  combined_skin_t_em_ifnghi$Subject
combined_skin_t_em_ifnghi$label <-
  combined_skin_t_em_ifnghi$clone_type
combined_skin_t_em_ifnghi$cell_type <- "CD8+ GZMK+ IFNGhi T Cells"

# we want the singletons to be the "reference" cell type
combined_skin_t_em_ifnghi$label <-
  relevel(combined_skin_t_em_ifnghi$label, ref = "Singleton")

de_skin_t_em_ifnghi_clone_type <-
  pseudobulk_DE(combined_skin_t_em_ifnghi, de_test = "pseudobulk_DESeq2")

saveRDS(de_skin_t_em_ifnghi_clone_type,
        file.path(path_objs, "downstream", "differential_expression",
                  "deseq2_combined_skin_t_em_ifnghi_clone_type.rds"))
```

# Total counts {.tabset}

Total counts by cell subset:

## Skin

```{r de-skin-counts-bar}
#| fig.dim = c(14, 8)

skin_no_de <-
  setdiff(unique(combined_skin$annotated_clusters),
          unique(de_genes_skin$cell_type))

plot_counts_cluster(clusters = de_genes_skin$cell_type,
                    tissue_type = "Skin",
                    clrs_specific = colors_annotated_skin,
                    x_axis = "Cell Type",
                    add_zeroes = skin_no_de,
                    details = "Differentially Expressed Genes")
```

## Skin (Recombined)

```{r de-skin-recombined-t-counts-bar}
#| fig.dim = c(14, 8)

skin_recombined_t_no_de <-
  setdiff(unique(combined_skin_recombined_t$annotated_clusters_expanded),
          unique(de_genes_skin_recombined_t$cell_type))

plot_counts_cluster(clusters = de_genes_skin_recombined_t$cell_type,
                    tissue_type = "Skin (Recombined T)",
                    clrs_specific = colors_annotated_skin_recombined_t,
                    x_axis = "Cell Type",
                    add_zeroes = skin_recombined_t_no_de,
                    details = "Differentially Expressed Genes")
```

```{r, eval = FALSE}
# shows everything
UpSetR::upset(fromList(split(de_genes_skin_recombined_t$gene,
                             de_genes_skin_recombined_t$cell_type)),
              nsets = n_distinct(de_genes_skin_recombined_t$cell_type),
              nintersects = NA, order.by = "freq",
              mainbar.y.label = "DEG Intersections")

# just the immune cell types
UpSetR::upset(fromList(split(de_genes_skin_recombined_t$gene,
                             de_genes_skin_recombined_t$cell_type)),
              nintersects = NA,
              sets = c("B Cells", "CD4- CD8- T Cells",
                       "CD4+ Tissue-Resident Memory T Cells",
                       "CD8+ GZMK+ IFNGhi T Cells",
                       "Cycling Dendritic Cells", "Langerhans", "Macrophages",
                       "Natural Killers"),
              # keep.order = TRUE,
              order.by = "freq",
              mainbar.y.label = "DEG Intersections (Immune Only)",
              text.scale = 1.1)
```

## Skin T Cells

```{r de-skin-t-counts-bar}
#| fig.dim = c(12, 8)

skin_t_no_de <-
  setdiff(unique(combined_skin_t$annotated_subclusters),
          unique(de_genes_skin_t$cell_type))

plot_counts_cluster(clusters = de_genes_skin_t$cell_type,
                    tissue_type = "Skin T Cells",
                    clrs_specific = colors_annotated_skin_t,
                    x_axis = "Cell Type",
                    add_zeroes = skin_t_no_de,
                    details = "Differentially Expressed Genes")
```

------------------------------------------------------------------------

# Heatmaps

Chemokines:

* [A guide to chemokines and their receptors (Hughes and Nibbs, 2018)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6120486/)
  * "Although chemokines were originally named according to specific functions, a systematic nomenclature was introduced in 2000 that includes a subfamily designation (i.e., CC, CXC, CX3C, or XC), followed by the letter L (denoting 'ligand'), and then a number according to when the gene was first isolated"
  * "There are 10 CCRs, 6 CXCRs, and a single CX3CR and XCR."
  * "Receptor specificity is complex: many chemokines bind to multiple cCKRs, and some cCKRs have many ligands"

## Skin {.tabset}

### All chemokine ligands {.tabset}

#### Full list

```{r de-skin-genes-chemokines-ligands-genes}
specific_gene <- "Chemokine Ligands"
genes_list <- c(str_c("CCL", 1:28), # from Hailong
                grep("^CX3CL", rownames(combined_skin), value = TRUE),
                str_c("CXCL", 1:17), # from Hailong
                grep("^XCL", rownames(combined_skin), value = TRUE))
```

```{r de-skin-genes-chemokines-ligands-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-chemokines-sig-ligands-genes}
specific_gene <- "Chemokine Ligands"
genes_list <- str(unique(de_genes_skin$gene),
                         "^CCL|^CX3CL|^CXCL|^XCL")
```

```{r de-skin-genes-chemokines-sig-ligands-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### All chemokine receptors {.tabset}

#### Full list

```{r de-skin-genes-chemokines-receptors-genes}
specific_gene <- "Chemokine Receptors"
genes_list <- grep("^(CCR|CX3CR|CXCR|XCR)", rownames(combined_skin),
                   value = TRUE)
```

```{r de-skin-genes-chemokines-receptors-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-chemokines-sig-receptors-genes}
specific_gene <- "Chemokine Receptors"
genes_list <- str(unique(de_genes_skin$gene),
                         "^CCR|^CX3CR|^CXCR|^XCR")
```

```{r de-skin-genes-chemokines-sig-receptors-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CCL {.tabset}

#### Full list

```{r de-skin-genes-ccl-genes}
specific_gene <- "CCL"
genes_list <- str_c(specific_gene, 1:28) # from Hailong
```

```{r de-skin-genes-ccl-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-ccl-sig-genes}
specific_gene <- "CCL"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-ccl-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CCR {.tabset}

#### Full list

```{r de-skin-genes-ccr-genes}
specific_gene <- "CCR"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin),
                   value = TRUE)
```

```{r de-skin-genes-ccr-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-ccr-sig-genes}
specific_gene <- "CCR"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-ccr-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CX3CL {.tabset}

#### Significant only

```{r de-skin-genes-cx3cl-sig-genes}
specific_gene <- "CX3CL"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-cx3cl-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CX3CR {.tabset}

#### Full list

```{r de-skin-genes-cx3cr-genes}
specific_gene <- "CX3CR"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin),
                   value = TRUE)
```

```{r de-skin-genes-cx3cr-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-cx3cr-sig-genes}
specific_gene <- "CX3CR"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-cx3cr-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CXCL {.tabset}

#### Full list

```{r de-skin-genes-cxcl-genes}
specific_gene <- "CXCL"
genes_list <- str_c(specific_gene, 1:17) # from Hailong
```

```{r de-skin-genes-cxcl-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-cxcl-sig-genes}
specific_gene <- "CXCL"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-cxcl-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CXCR {.tabset}

#### Full list

```{r de-skin-genes-cxcr-genes}
specific_gene <- "CXCR"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin),
                   value = TRUE)
```

```{r de-skin-genes-cxcr-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-cxcr-sig-genes}
specific_gene <- "CXCR"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-cxcr-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### GBP {.tabset}

> Guanylate-binding proteins (GBP) are a family of dynamin-related large GTPases which are expressed in response to interferons and other pro-inflammatory cytokines. GBPs mediate a broad spectrum of innate immune functions against intracellular pathogens ranging from viruses to bacteria and protozoa. - https://pubmed.ncbi.nlm.nih.gov/29174633/

> Interferon-gamma (IFN-γ) is a pleiotropic cytokine that exerts important functions in inflammation, infectious diseases, and cancer. The large GTPase human guanylate-binding protein 1 (GBP-1) is among the most strongly IFN-γ-induced cellular proteins. - https://pmc.ncbi.nlm.nih.gov/articles/PMC6156764/

#### Full list

```{r de-skin-genes-gbp-genes}
specific_gene <- "GBP"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin),
                   value = TRUE)
```

```{r de-skin-genes-gbp-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-gbp-sig-genes}
specific_gene <- "GBP"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-gbp-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### IFN

```{r de-skin-genes-ifn-genes}
specific_gene <- "IFN"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-ifn-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### IFNG

```{r de-skin-genes-ifng-genes}
specific_gene <- "IFNG"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-ifng-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### IL {.tabset}

#### Full list

```{r de-skin-genes-il-genes}
specific_gene <- "IL"

# from Hailong
# "Interleukin 30 (IL-30) forms one chain of the heterodimeric cytokine called interleukin 27 (IL-27)"
genes_list <- c(paste0("IL", 1:39),
                "IL1A", "IL1B", "IL12A", "IL12B", "IL17A", "IL17B", "IL23A",
                "CXCL8", # CXCL8 is IL8
                "TXLNA", # TXLNA is IL14
                "IL28A", # IFNL2 is IL28A
                "IFNL3", # IFNL3 is IL28B
                "IFNL1", # IFNL1 is IL29
                "IL36A")
```

```{r de-skin-genes-il-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-il-sig-genes}
specific_gene <- "IL"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-il-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### ISGs

```{r de-skin-genes-isgs-genes}
specific_gene <- "Interferon-Stimulated"
genes_list <- c("CCL5", "CCL7", "IL10", "IL32", "IRF1", "IRF9", "ISG20",
                "SOCS3", "STAT1", "STAT2", "TAPBP", "TNFRSF14")
```

```{r de-skin-genes-isgs-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### Other {.tabset}

```{r de-skin-genes-other-genes}
specific_gene <- "Other"
genes_list <- c("LCE2A", "LCE2B", "LCE3A", "LCE3B") # from Hailong
```

```{r de-skin-genes-other-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### MMP {.tabset}

#### Full list

```{r de-skin-genes-mmp-genes}
specific_gene <- "MMP"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin),
                   value = TRUE)
```

```{r de-skin-genes-mmp-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-mmp-sig-genes}
specific_gene <- "MMP"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-mmp-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### TLR {.tabset}

#### Full list

```{r de-skin-genes-tlr-genes}
specific_gene <- "TLR"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin),
                   value = TRUE)
```

```{r de-skin-genes-tlr-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-tlr-sig-genes}
specific_gene <- "TLR"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-tlr-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### TNF

```{r de-skin-genes-tnf-genes}
specific_gene <- "TNF"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-tnf-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### XCL {.tabset}

#### Full list

```{r de-skin-genes-xcl-genes}
specific_gene <- "XCL"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin),
                   value = TRUE)
```

```{r de-skin-genes-xcl-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-xcl-sig-genes}
specific_gene <- "XCL"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-xcl-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### XCR {.tabset}

#### Full list

```{r de-skin-genes-xcr-genes}
specific_gene <- "XCR"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin),
                   value = TRUE)
```

```{r de-skin-genes-xcr-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-genes-xcr-sig-genes}
specific_gene <- "XCR"
genes_list <- str_subset(string = unique(de_genes_skin$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-genes-xcr-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

## Skin (Recombined) {.tabset}

### All chemokine ligands {.tabset}

#### Full list

```{r de-skin-recombined-genes-chemokines-ligands-genes}
specific_gene <- "Chemokine Ligands"
# doing it Hailong's way e.g. str_c("CCL", 1:28) misses genes like CCL3L1
genes_list <- c(str_c("CCL", 1:28), # from Hailong
                grep("^CX3CL", rownames(combined_skin_recombined_t), value = TRUE),
                str_c("CXCL", 1:17), # from Hailong
                grep("^XCL", rownames(combined_skin_recombined_t), value = TRUE))
```

```{r de-skin-recombined-genes-chemokines-ligands-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-chemokines-sig-ligands-genes}
specific_gene <- "Chemokine Ligands"
genes_list <- str(unique(de_genes_skin_recombined_t$gene),
                         "^CCL|^CX3CL|^CXCL|^XCL")
```

```{r de-skin-recombined-genes-chemokines-sig-ligands-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### All chemokine receptors {.tabset}

#### Full list

```{r de-skin-recombined-genes-chemokines-receptors-genes}
specific_gene <- "Chemokine Receptors"
genes_list <- grep("^(CCR|CX3CR|CXCR|XCR)", rownames(combined_skin_recombined_t),
                   value = TRUE)
```

```{r de-skin-recombined-genes-chemokines-receptors-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-chemokines-sig-receptors-genes}
specific_gene <- "Chemokine Receptors"
genes_list <- str(unique(de_genes_skin_recombined_t$gene),
                         "^CCR|^CX3CR|^CXCR|^XCR")
```

```{r de-skin-recombined-genes-chemokines-sig-receptors-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```


### CCL {.tabset}

#### Full list

```{r de-skin-recombined-genes-ccl-genes}
specific_gene <- "CCL"
genes_list <- str_c(specific_gene, 1:28) # from Hailong
```

```{r de-skin-recombined-genes-ccl-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-ccl-sig-genes}
specific_gene <- "CCL"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-ccl-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CCR {.tabset}

#### Full list

```{r de-skin-recombined-genes-ccr-genes}
specific_gene <- "CCR"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin_recombined_t),
                   value = TRUE)
```

```{r de-skin-recombined-genes-ccr-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-ccr-sig-genes}
specific_gene <- "CCR"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-ccr-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CX3CL {.tabset}

#### Significant only

```{r de-skin-recombined-genes-cx3cl-sig-genes}
specific_gene <- "CX3CL"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-cx3cl-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CX3CR {.tabset}

#### Full list

```{r de-skin-recombined-genes-cx3cr-genes}
specific_gene <- "CX3CR"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin_recombined_t),
                   value = TRUE)
```

```{r de-skin-recombined-genes-cx3cr-bar}
#| fig.dim = c(10, max(6, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-cx3cr-sig-genes}
specific_gene <- "CX3CR"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-cx3cr-sig-bar}
#| fig.dim = c(10, max(6, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CXCL {.tabset}

#### Full list

```{r de-skin-recombined-genes-cxcl-genes}
specific_gene <- "CXCL"
genes_list <- str_c(specific_gene, 1:17) # from Hailong
```

```{r de-skin-recombined-genes-cxcl-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-cxcl-sig-genes}
specific_gene <- "CXCL"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-cxcl-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CXCR {.tabset}

#### Full list

```{r de-skin-recombined-genes-cxcr-genes}
specific_gene <- "CXCR"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin_recombined_t),
                   value = TRUE)
```

```{r de-skin-recombined-genes-cxcr-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-cxcr-sig-genes}
specific_gene <- "CXCR"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-cxcr-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### GBP {.tabset}

> Guanylate-binding proteins (GBP) are a family of dynamin-related large GTPases which are expressed in response to interferons and other pro-inflammatory cytokines. GBPs mediate a broad spectrum of innate immune functions against intracellular pathogens ranging from viruses to bacteria and protozoa. - https://pubmed.ncbi.nlm.nih.gov/29174633/

> Interferon-gamma (IFN-γ) is a pleiotropic cytokine that exerts important functions in inflammation, infectious diseases, and cancer. The large GTPase human guanylate-binding protein 1 (GBP-1) is among the most strongly IFN-γ-induced cellular proteins. - https://pmc.ncbi.nlm.nih.gov/articles/PMC6156764/

#### Full list

```{r de-skin-recombined-genes-gbp-genes}
specific_gene <- "GBP"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin_recombined_t),
                   value = TRUE)
```

```{r de-skin-recombined-genes-gbp-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-gbp-sig-genes}
specific_gene <- "GBP"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-gbp-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### HLA {.tabset}

#### Full list

```{r de-skin-recombined-genes-hla-genes}
specific_gene <- "HLA"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin_recombined_t),
                   value = TRUE)
```

```{r de-skin-recombined-genes-hla-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-hla-sig-genes}
specific_gene <- "HLA"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-hla-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### IFN {.tabset}

#### Full list

```{r de-skin-recombined-genes-ifn-genes}
specific_gene <- "IFN"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin_recombined_t),
                   value = TRUE)
```

```{r de-skin-recombined-genes-ifn-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-ifn-sig-genes}
specific_gene <- "IFN"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-ifn-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### IL {.tabset}

#### Full list

```{r de-skin-recombined-genes-il-genes}
specific_gene <- "IL"

# from Hailong
# "Interleukin 30 (IL-30) forms one chain of the heterodimeric cytokine called interleukin 27 (IL-27)"
genes_list <- c(paste0("IL", 1:39),
                "IL1A", "IL1B", "IL12A", "IL12B", "IL17A", "IL17B", "IL23A",
                "CXCL8", # CXCL8 is IL8
                "TXLNA", # TXLNA is IL14
                "IL28A", # IFNL2 is IL28A
                "IFNL3", # IFNL3 is IL28B
                "IFNL1", # IFNL1 is IL29
                "IL36A")
```

```{r de-skin-recombined-genes-il-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-il-sig-genes}
specific_gene <- "IL"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-il-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### ISGs

```{r de-skin-recombined-genes-isgs-genes}
specific_gene <- "Interferon-Stimulated"
genes_list <- c("CCL5", "CCL7", "IL10", "IL32", "IRF1", "IRF9", "ISG20",
                "SOCS3", "STAT1", "STAT2", "TAPBP", "TNFRSF14")
```

```{r de-skin-recombined-genes-isgs-bar}
#| fig.dim = c(12, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### Other {.tabset}

```{r de-skin-recombined-genes-other-genes}
specific_gene <- "Other"
genes_list <- c("LCE2A", "LCE2B", "LCE3A", "LCE3B") # from Hailong
```

```{r de-skin-recombined-genes-other-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### MMP {.tabset}

#### Full list

```{r de-skin-recombined-genes-mmp-genes}
specific_gene <- "MMP"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin_recombined_t),
                   value = TRUE)
```

```{r de-skin-recombined-genes-mmp-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-mmp-sig-genes}
specific_gene <- "MMP"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-mmp-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### TLR {.tabset}

#### Full list

```{r de-skin-recombined-genes-tlr-genes}
specific_gene <- "TLR"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin_recombined_t),
                   value = TRUE)
```

```{r de-skin-recombined-genes-tlr-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-tlr-sig-genes}
specific_gene <- "TLR"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-tlr-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### TNF

```{r de-skin-recombined-genes-tnf-genes}
specific_gene <- "TNF"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-tnf-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### XCL {.tabset}

#### Full list

```{r de-skin-recombined-genes-xcl-genes}
specific_gene <- "XCL"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin_recombined_t),
                   value = TRUE)
```

```{r de-skin-recombined-genes-xcl-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-xcl-sig-genes}
specific_gene <- "XCL"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-xcl-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### XCR {.tabset}

#### Full list

```{r de-skin-recombined-genes-xcr-genes}
specific_gene <- "XCR"
genes_list <- grep(paste0("^", specific_gene), rownames(combined_skin_recombined_t),
                   value = TRUE)
```

```{r de-skin-recombined-genes-xcr-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

#### Significant only

```{r de-skin-recombined-genes-xcr-sig-genes}
specific_gene <- "XCR"
genes_list <- str_subset(string = unique(de_genes_skin_recombined_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-recombined-genes-xcr-sig-bar}
#| fig.dim = c(10, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_recombined_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin (Recombined)"),
                 x_angle = 90, col_size = 12)
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

## Skin T Cells {.tabset}

*All significant only except for the ISGs.*

### CCL

```{r de-skin-t-genes-ccl-sig-genes}
specific_gene <- "CCL"
genes_list <- str_subset(string = unique(de_genes_skin_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-t-genes-ccl-sig-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CCR

```{r de-skin-t-genes-ccr-sig-genes}
specific_gene <- "CCR"
genes_list <- str_subset(string = unique(de_genes_skin_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-t-genes-ccr-sig-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CX3CL

```{r de-skin-t-genes-cx3cl-sig-genes}
specific_gene <- "CX3CL"
genes_list <- str_subset(string = unique(de_genes_skin_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-t-genes-cx3cl-sig-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CX3CR

```{r de-skin-t-genes-cx3cr-sig-genes}
specific_gene <- "CX3CR"
genes_list <- str_subset(string = unique(de_genes_skin_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-t-genes-cx3cr-sig-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CXCL

```{r de-skin-t-genes-cxcl-sig-genes}
specific_gene <- "CXCL"
genes_list <- str_subset(string = unique(de_genes_skin_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-t-genes-cxcl-sig-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### CXCR

```{r de-skin-t-genes-cxcr-sig-genes}
specific_gene <- "CXCR"
genes_list <- str_subset(string = unique(de_genes_skin_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-t-genes-cxcr-sig-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### IFITM

```{r de-skin-t-genes-ifitm-sig-genes}
specific_gene <- "IFITM"
genes_list <- str_subset(string = unique(de_genes_skin_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-t-genes-ifitm-sig-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### IFN

```{r de-skin-t-genes-ifn-sig-genes}
specific_gene <- "IFN"
genes_list <- str_subset(string = unique(de_genes_skin_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-t-genes-ifn-sig-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### IFNG

```{r de-skin-t-genes-ifng-sig-genes}
specific_gene <- "IFNG"
genes_list <- str_subset(string = unique(de_genes_skin_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-t-genes-ifng-sig-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### IL

```{r de-skin-t-genes-il-sig-genes}
specific_gene <- "IL"
genes_list <- str_subset(string = unique(de_genes_skin_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-t-genes-il-sig-bar}
#| fig.dim = c(14, max(4, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### ISGs

```{r de-skin-t-genes-isgs-genes}
specific_gene <- "Interferon-Stimulated"
genes_list <- c("CCL5", "CCL7", "IL10", "IL32", "IRF1", "IRF9", "ISG20",
                "SOCS3", "STAT1", "STAT2", "TAPBP", "TNFRSF14")
```

```{r de-skin-t-genes-isgs-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### XCL

```{r de-skin-t-genes-xcl-sig-genes}
specific_gene <- "XCL"
genes_list <- str_subset(string = unique(de_genes_skin_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-t-genes-xcl-sig-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```

### XCR

```{r de-skin-t-genes-xcr-sig-genes}
specific_gene <- "XCR"
genes_list <- str_subset(string = unique(de_genes_skin_t$gene),
                         pattern = paste0("^", specific_gene))
```

```{r de-skin-t-genes-xcr-sig-bar}
#| fig.dim = c(12, max(3, length(genes_list)/2 + 2))

if (length(genes_list) > 0) {
  heatmap_DESeq2(DESeq2_out = de_skin_t,
                 gene_list = genes_list,
                 FDR_cutoff = fdr_cutoff,
                 title = paste("Gene Expression of", specific_gene,
                               "in the Skin T Cells"))
} else {
  cat(paste0("No genes found for ", specific_gene, "."))
}
```
